Economic Evaluation of Geographically Weighted Regression Analysis for Site-specific Nitrogen Management

### 2023-02-13 Writing/manuscript_gwr.rmd
manuscript for JAAEA

**Abstract**: Geographically weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to derive site-specific input rate recommendations from data generated by on-farm precision experimentation (OFPE). The key objectives of this paper are two-fold. The first objective is to evaluate economically the two sets of functional-form specifications used by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that the relative performance of the two approaches vary depending on the crop-fertilizer price ratios due to their tendency to over-/under-estimate EONR (**perhaps use the full name of EONR here**). The second and more important, we investigate the possibility of bias in estimating the economic value of site-specific recommendation maps based on GWR. We show that both GWR approaches suffer significant positive bias, which can lead to large over-estimation of the economic value of applying GWR methods. Our study is the first to systematically examine the economic (not statistical) properties of the use of GWR in OPFE settings. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management.

@@

(**I feel we are talking too much about the soil sampling data; maybe we can make this paragraph more concise**) On any given field, it is typical for U.S. Midwestern row crop farmers to pay for 1-ha grid soil sampling and chemical analysis every three or four years [@morris2018strengths]. The data is often used to derive fertilizer application recommendations, for example by applying comparatively simple formulas provided by university extension services (e.g., Culman, et al. 2020). But the expense of soil sampling and analysis is driving development of sensor technologies that gather data on site-specific field and crop characteristics or proxies of those characteristics. For example, Trimble Agriculture's GreenSeeker and AgLeader's OptRx® systems use optical sensors mounted on fertilizer application booms to digitally record vegetative index imagery as farm machinery is driven through fields during early crop growth stages. GreenSeeker then basically uses Oklahoma State University's Sensor-based N Rate Calculator to convert those data into recommended nitrogen fertilizer (N) application rate maps ($R_xs$). Veris Technologies manufactures sensor equipment that can be towed through fields to gather apparent electroconductivity (EC) data, which can be used to infer properties of soil content. The "decision tool" software FieldRevealTM allows users to upload EC data, elevation data, and vegetative index data, which FieldRevealTM's professional agronomists analyze to provide $R_x$s to farmers. FieldApex creates N $R_x$s using foliar vegetative index data from satellite imagery, and SlantRange uses drone imagery for similar purposes.

@@

This study has two key objectives. The first is to determine which of the model specifications used by @rakshit2020 and @trevisan2021spatial result in more profitable $R_xs$. While both @rakshit2020 and @trevisan2021spatial proposed using GWR to estimate site-specific yield response functions, they used different model specifications. Both assumed a quadratic response of yield to nitrogen rate ($N$). However, while @trevisan2021spatial estimated the coefficient on $N^2$ as a global parameter, @rakshit2020 estimated it locally for each point. At first, it appears that the specification used by @rakshit2020 is superior since it is more flexible than the specification used by @trevisan2021spatial. However, the flexibility can be a disadvantage rather than an advantage because the estimation of EONR is highly sensitive to the coefficient estimates on $N^2$, and GWR often provides inaccurate local estimation of that coefficient.

@@

Figure \@ref(fig:field-layout) shows the simulated field's layout and defines the spatial units of the simulated on-farm precision experiments in this study. The basic spatial unit of the field is a 6m $\times$ 6m "cell", which is the resolution at which the field characteristics data (i.e., parameters of the "true" yield response function) were allowed to vary. It was assumed, however, that yields could not be monitored at such a high resolution. Rather, the mechanical harvester was assumed to be 18m wide, and, in cases in which the N rate was unchanged, in need of harvesting on a run of no less than 12m to provide accurate yield data. Therefore, the yield data's resolution was 18m $\times$ 12m (3 cells $\times$ 2 cells, called a "subplot"). The N rate unit consisted of five (**it should be six**) "subplots", that is, 18m $\times$ 72m (3 cells $\times$ 12 cells, called a "plot") to accommodate the variable-rate applicator's application resolution. A 12m "transition zone" was placed between two plots (that is, a 6m "transition zone" on each end of a plot), so that when the experimental N rate changed, the harvester had 12m to "clear" and begin to record accurately on a plot with an N rate different from its predecessor's.

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6-m-wide (**is it wide or long?**) headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12-m-wide "transition zones" were not included in the data analyzed. In total there were 1440 subplots available for experimental data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $k \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet $(j, k)$ is used to identify individual cells.

>>>>

**Abstract**: Geographically weighted regression (GWR) has been presented as a prominent tool to estimate site-specific yield response functions to derive site-specific input rate recommendations from data generated by on-farm precision experimentation (OFPE). The key objectives of this paper are two-fold. The first is to economically evaluate the two sets of functional-form specifications used by two recent published studies of site-specific input management that apply GWR to data from OFPE. We show that the relative performance of the two approaches vary depending on the crop-fertilizer price ratios due to their tendency to over-/under-estimate the economically optimal nitrogen rate. The second and more important, we investigate the possibility of bias in estimating the economic value of site-specific recommendation maps based on GWR. We show that both GWR approaches suffer significant positive bias, which can lead to large over-estimation of the economic value of applying GWR methods. Our study is the first to systematically examine the economic (not statistical) properties of the use of GWR in OPFE settings. We reveal that in many circumstances using GWR is far less profitable than claimed by previous studies, which provides a necessary cautionary message to researchers who might otherwise “over-sell” to other researchers the value of using GWR methods in empirical work involving the estimation of site-specific input management.

@@

On any given field, it is typical for U.S. Midwestern row crop farmers to pay for 1-ha grid soil sampling and chemical analysis every three or four years [@morris2018strengths]. The data is often used to derive fertilizer application recommendations, for example by applying comparatively simple formulas provided by university extension services (e.g., Culman, et al. 2020). But the expense of soil sampling and analysis is driving development of sensor technologies that gather data on site-specific field and crop characteristics or proxies of those characteristics. For example, Trimble Agriculture's GreenSeeker and AgLeader's OptRx® systems use optical sensors mounted on fertilizer application booms to digitally record vegetative index imagery as farm machinery is driven through fields during early crop growth stages. GreenSeeker then basically uses Oklahoma State University's Sensor-based N Rate Calculator to convert those data into recommended nitrogen fertilizer (N) application rate maps ($R_xs$). Veris Technologies manufactures sensor equipment that can be towed through fields to gather apparent electroconductivity (EC) data, which can be used to infer properties of soil content. The "decision tool" software FieldRevealTM allows users to upload EC data, elevation data, and vegetative index data, which FieldRevealTM's professional agronomists analyze to provide $R_x$s to farmers. FieldApex creates N $R_x$s using foliar vegetative index data from satellite imagery, and SlantRange uses drone imagery for similar purposes.

@@

This study has two key objectives. The first is to determine which of the model specifications used by @rakshit2020 and @trevisan2021spatial result in more profitable $R_xs$. While both @rakshit2020 and @trevisan2021spatial proposed using GWR to estimate site-specific yield response functions, they used different model specifications. Both assumed a quadratic response of yield to nitrogen rate ($N$). However, while @trevisan2021spatial estimated the coefficient on $N^2$ as a global parameter, @rakshit2020 estimated it locally for each point. At first, it appears that the specification used by @rakshit2020 is superior since it is more flexible than the specification used by @trevisan2021spatial. However, the flexibility can be a disadvantage rather than an advantage because the estimation of economically optimal nitrogen rate (EONR) is highly sensitive to the coefficient estimates on $N^2$, and GWR often provides inaccurate local estimation of that coefficient.

@@

Figure \@ref(fig:field-layout) shows the simulated field's layout and defines the spatial units of the simulated on-farm precision experiments in this study. The basic spatial unit of the field is a 6m $\times$ 6m "cell", which is the resolution at which the field characteristics data (i.e., parameters of the "true" yield response function) were allowed to vary. It was assumed, however, that yields could not be monitored at such a high resolution. Rather, the mechanical harvester was assumed to be 18m wide, and, in cases in which the N rate was unchanged, in need of harvesting on a run of no less than 12m to provide accurate yield data. Therefore, the yield data's resolution was 18m $\times$ 12m (3 cells $\times$ 2 cells, called a "subplot"). The N rate unit consisted of five "subplots", that is, 18m $\times$ 72m (3 cells $\times$ 12 cells, called a "plot") to accommodate the variable-rate applicator's application resolution. A 12m "transition zone" was placed between two plots (that is, a 6m "transition zone" on each end of a plot), so that when the experimental N rate changed, the harvester had 12m to "clear" and begin to record accurately on a plot with an N rate different from its predecessor's.

The field was 432m (72 cells) wide and 864m (144 cells) long, making its area 37.3 ha. Data from the 6-m-long headlands, side-lands (there are no transition zones on the north and south ends of the field), and 12-m-wide "transition zones" were not included in the data analyzed. In total there were 1440 subplots available for experimental data analysis. The index $j \in {1, 2, \dots, 1440}$ identifies the field's subplots, and the index $k \in {1, \dots , 6}$ identifies the cells within a subplot. The couplet $(j, k)$ is used to identify individual cells.
